---
title: "Exploratory Data Visualizations - Federal Budget Data"
author: "Ian Lyons"
date: "1/27/2019"
output: 
  html_document: 
    fig_caption: yes
    fig_height: 8.5
    fig_width: 11
    code_folding: hide
urlcolor: blue

---

```{r setup, echo=FALSE, include=FALSE}
library(tidyverse)
library(rmarkdown)
library(here)
library(gghighlight)
library(ggpubr)
library(formatR)
library(tinytex)
library(scales)
```

## Read in Tax Expenditure and Federal Outlays Data

* Source for tax expenditures on the Treasury Department Website: 
https://home.treasury.gov/policy-issues/tax-policy/tax-expenditures

* Source for federal outlays on the Office of Management and Budget website: 
https://www.whitehouse.gov/omb/supplemental-materials/
```{r Read in Data, include=FALSE, echo=FALSE}
base_df <- read_csv(file = 'raw_data/tax_expenditures_18to28_renamed.csv')
rev <- read_csv(file = 'raw_data/receipts-fy2019.csv')
budg_2019 <- read_csv(file = 'raw_data/outlays-fy2019.csv')
```


## Reshape Tax Expenditure Dataframe For Plot 1

* Create a long-form version of the tax expenditures dataset to use in a time-series plot. That is, create one row for each category, detail, and fiscal year. 
* Drop the cumulative 'FY2019_to_2028' column since I won't be using it.

```{r Reshape tax expenditures data}
tax_long <- gather(base_df, fiscal_year, expenditure, '2018':'2028', factor_key=TRUE)
tax_long$FY2019_to_2028 <- NULL
```


## Plot 1: Largest Tax Expenditures By  Category
Tax expenditures are what we usually refer to as tax loopholes or tax breaks. The US Department of Treasury classifies these tax provisions as tax "expenditures" in recognition of the fact that they are "often viewed as alternatives to other policy instruments, such as spending or regulatory programs" ([Treasury](https://home.treasury.gov/policy-issues/tax-policy/tax-expenditures)). 

```{r Top 7 2018 Tax Expenditures}
time_series <- tax_long %>% select(fiscal_year, Category, expenditure) %>% filter(fiscal_year %in% c(2018)) %>% 
  group_by(fiscal_year, Category) %>% tally(expenditure)

time_series <- mutate(.data = time_series, amount_billions = n/1000)
time_series$n <- NULL
time_series <- arrange(time_series, -amount_billions)
tax_col_names <- c("Aid to State & Local Govts", "Commerce & Housing", "Education, Training, Employment, and Social Services", "Fiscal Assistance", "Health", "Income Security", "International Affairs")
top7 <- top_n(time_series, 7)


bar_to_plot <- ggplot(data=top7, aes(x=tax_col_names, y=amount_billions, color=tax_col_names, fill=tax_col_names)) + 
  geom_col(position = "dodge", color="black")

plot1 <- bar_to_plot + 
    scale_y_continuous(expand = c(0,0), name = "Tax Expenditures in Billions of Dollars", labels = scales::dollar_format(accuracy = 1, suffix = 'B')) + 
    scale_x_discrete() + 
    scale_fill_brewer(palette="Set2") +
  
    labs(title = "The United States spends trillions of dollars on tax loopholes each year", 
         subtitle = "US Federal Government Tax Expenditures In 2018",
         caption = "Source: United States. Department of Treasury. Tax Policy: Tax Expenditures. \n * These estimates are made relative to current law as of July 1, 2018", 
         x = "Category of Tax Break") +              
            
          theme(plot.title = element_text(color="black", size=22, face="bold", hjust = 0.5),
              plot.subtitle = element_text(color="black", size=18, hjust =0.5),
              axis.title.x = element_text(color="black", size=18), axis.text.x = element_blank(),
              axis.title.y = element_text(color="black", size=18),
              plot.caption = element_text(color="black", size=12, face="italic")
              )
plot1
```

<br><br>


## The largest individual tax expenditures
```{r Top Tax Expenditures}
taxex <- filter(tax_long, fiscal_year == 2018)
top10 <- top_n(taxex, 10)
top10 <- arrange(top10, -expenditure)

theme1 <- theme(plot.title = element_text(color="black", size=22, face="bold", hjust = 0.5),
              plot.subtitle = element_text(color="black", size=18, hjust =0.5),
              axis.title.x = element_text(color="black", size=18), axis.text.x = element_blank(),
              axis.title.y = element_text(color="black", size=18),
              plot.caption = element_text(color="black", size=12, face="italic"),
              legend.position="none")

descr <- labs(title = "US Federal Government Tax Expenditures, 2018-2028", 
              subtitle = "The United States spends trillions of dollars on tax loopholes each year",
              caption = "Source: United States. Department of Treasury. Tax Policy: Tax Expenditures. \n 
              * These estimates are made relative to current law as of July 1, 2018", 
              x = "Description of Tax Break")

expend_plot <- ggplot(data=top10, aes(x=Details, y=expenditure, color=Details, fill=Details)) + 
  geom_col(color="black") + coord_flip() +
  scale_y_continuous(expand = c(0,0), name = "Tax Expenditures in Billions of Nominal Dollars", 
                labels = scales::dollar_format(accuracy = 1, suffix = 'B')) + 
  scale_x_discrete(labels = scales::wrap_format(18)) + 
  scale_fill_brewer(palette="Set2") +
  descr +
  theme1

expend_plot
```


# Prep For Plots 2 & 3: Group by agency name and sum expenditures within the same agency and year.

* Create a long-form version of the outlays dataset to use in a time-series plot. That is, create one row for each line item (agency name, bureau name, account name, etc.) and fiscal year. 

```{r}
budget_longform <- gather(budg_2019, fiscal_year, expenditure, '1962':'2018', factor_key=TRUE)

## Keep only descriptive column names rather than codes. 
keep_cols <- c('Agency Name', 'Bureau Name', 'Account Name', 'Treasury Agency Code', 'Subfunction Title', 'BEA Category', 'Grant/non-grant split' , 'On- or Off- Budget', 'fiscal_year', 'expenditure')

budget_longform <- budget_longform[keep_cols]
```

* Take the reshaped longform budget data and group by fiscal year and agency name. 

```{r}
outlays <- budget_longform %>% select(fiscal_year, `Agency Name`, expenditure) %>%
			group_by(fiscal_year, `Agency Name`) %>%
			tally(expenditure)

## Give the expenditures column a descriptive name			
outlays <- mutate(.data = outlays, dollars_thousands = n)
outlays$n <- NULL

## Rename `Agency Name` in snake case.
colnames(outlays)[2] <- 'agency_name'

```

<br><br>

## Plot 2: FacetWrap Grid Showing the Annual Outlays 1995-2018 of Some of the Largest Agencies & Departments

* Filter to the fiscal years 1995 to 2018.
* Filter to only the 12 agencies/departments with the largest budgets.

```{r Facetwrap 1995-2018}
outlays %>%
  filter(fiscal_year  %in% c(1995:2018)) %>%
  filter(agency_name %in% c("Department of Agriculture",
  "Department of Defense--Military Programs",
  "Department of Education", 
  "Department of Health and Human Services",
  "Department of Homeland Security",
  "Department of Housing and Urban Development", "Department of Labor", "Department of the Treasury", "Department of Transportation", "Department of Veterans Affairs")) %>%
	      
  ggplot(aes(x = fiscal_year, y = dollars_thousands, fill = agency_name)) +
  geom_histogram(stat = "sum") + 
  facet_wrap(~agency_name) +
  scale_y_continuous(name = "Outlays (Thousands of US Dollars)", labels = scales::dollar_format(scale=.000001, accuracy = .01, suffix = 'B')) +

  labs(title = "Selected Departmental Outlays Since 1995",
       subtitle = "Some agencies, like the Dept of Defense and the Treasury, have experienced significant shifts in outlays since 1995.",
       caption="Source: United States. Office of Management and Budget. Public Budget Database.",
       x = "Fiscal Year",
       y = "Outlays In Thousands of US Dollars ") + 
       
  theme(axis.text.x = element_text(angle = 90),
        strip.text = element_text(face = "plain", 
        size = rel(0.9)),
        plot.title = element_text(color="black", size=14, face="bold", hjust = 0.5),
        plot.subtitle = element_text(color="black", size=12, hjust =0.5),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14),
        plot.caption = element_text(color="black", size=8, face="italic"),
        legend.position="none"
        )
```


<br><br>

# Plot 3: Federal Government Outlays by Agency By Year
* Use the same outlays dataset. 
* I take the federal government outlays dataset and choose the top 8 departments by total amount spent over the last 9 years.
* Limit the years to just the last 9 so that the point plot is comprehensible. 
* Group by agency name and fiscal year. 

```{r}
outlays %>%
  filter(agency_name %in% c("Social Security Administration","Department of Health and Human Services",
                            "Department of Defense--Military Programs", "Department of Agriculture", 
                            "Department of Veterans Affairs", "Department of Labor")) %>% 
	
	filter(fiscal_year %in% c(2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018)) %>%
	
  group_by(agency_name, fiscal_year) %>%
  summarize(dollars_thousands = sum(dollars_thousands)) %>%
  
      ggplot(aes(x = fiscal_year, y = (dollars_thousands), color = agency_name, group=agency_name)) +
      geom_point() + geom_line() + ggpubr::rotate_x_text() + theme(legend.position="none") +
      scale_y_log10(name = "Outlays (Billions)", labels = scales::dollar_format(scale=.000001, accuracy = 1, suffix = 'B')) +
        
        labs(title = "Since the Recession, Departments Have Had Drastically Different Treatment",
          subtitle = "Labor's stimulus funding was short-lived, and DoD's expenditures decrerased as wars winded down",
          caption = "Source: United States. Office of Management and Budget. Public Budget Database.",
          x = "Year",
          y = "Outlays (Thousands of US Dollars)" +              
            
          theme(plot.title = element_text(color="black", size=14, face="bold", hjust = 0.5),
              plot.subtitle = element_text(color="black", size=12, hjust =0.5),
              axis.title.x = element_text(color="black", size=14, angle = 90),
              axis.title.y = element_text(color="black", size=14),
              plot.caption = element_text(color="black", size=8, face="italic")
              )
        )
ggsave(filename = 'visuals/FederalOutlaysSince2010.pdf', width = 11, height = 8.5)
outlays
```



## Plot 4: Discretionary vs Mandatory at the Agency Level

I don't like this plot and I'm going to get rid of it. ¯\_(ツ)_/¯

```{r, echo=FALSE}
spendtype <- budget_longform %>% filter(fiscal_year %in% c(2018)) %>% filter(expenditure > 0) %>% select(fiscal_year, `Agency Name`, `BEA Category`, `Grant/non-grant split`, expenditure) %>%
    group_by(fiscal_year, `Agency Name`, `BEA Category`) %>%
    tally(expenditure)

v2_spendtype <- spread(spendtype, key = `BEA Category`, value = `n`)
colnames(v2_spendtype) <- c('fiscal_year', 'agency_name', 'discretionary', 'mandatory', 'net_interest')
v2_spendtype <- v2_spendtype %>% rowwise() %>% mutate(prop_mandatory = mandatory / sum(discretionary, mandatory, net_interest, na.rm=TRUE))
v2v2_spendtype <- v2_spendtype %>% rowwise() %>% mutate(total_spend = sum(discretionary, mandatory, net_interest, na.rm=TRUE))
v2_spendtype <- mutate(.data = v2_spendtype,  mandatory = coalesce(mandatory, 0))
v2_spendtype <- mutate(.data = v2_spendtype,  discretionary = coalesce(discretionary, 0))
v2_spendtype <- mutate(.data = v2_spendtype,  net_interest = coalesce(net_interest, 0))
v2_spendtype <- mutate(.data = v2_spendtype,  prop_mandatory = coalesce(prop_mandatory, 0))
v2_spendtype <- mutate(v2_spendtype, total_spend = discretionary +  mandatory + net_interest)

ggplot(data=v2_spendtype, aes(x = mandatory, y=discretionary)) + 
    geom_point(aes(size=total_spend, color=prop_mandatory)) + 
    scale_y_log10(name = "Mandatory Outlays (2018)", labels = scales::dollar_format(scale=.000001, accuracy = .01, suffix = 'B')) + 
    scale_x_log10(name = "Discretionary Outlays (2018)", labels = scales::dollar_format(scale=.000001, accuracy = .01, suffix = 'B')) + 
    
    labs(title = "Top Spending Agencies Have a Mix of Discretionary and Mandatory Spending",
         subtitle = "The DOD is one of the largest discretionary outlays, while HHS is one of the largerst mandatory outlays",
         caption = "Source: United States. Office of Management and Budget. Public Budget Database." +  
             theme(plot.title = element_text(color="black", size=18, face="bold", hjust = 0.5),
                   axis.title.x = element_text(color="black", size=18), axis.text.x = element_text(angle = 90, hjust = 1),
                   axis.title.y = element_text(color="black", size=18)))+ 
    
    geom_text(size=2, aes(label=ifelse(mandatory>0 & discretionary > 0, as.character(agency_name),''), hjust=0, vjust=0)) + 
    
    scale_colour_gradient2(low = "red", mid = "white", high = "blue", midpoint = .5, space = "Lab", na.value = "grey50", guide = "colourbar", aesthetics = "colour")
```



## Plot 5: Share of Spending Over Time  
The share of federal spending classified as mandatory has grown as a percentage of total spending since the 1960s. 

Mandatory spending is spending controlled by laws other than annual appropriations acts. This spending has been mandated by Congress in the past and does does have to be reappropriated each year. Most mandatory programs are fall under health care (e.g. Medicare, Medicaid), social security, and other income support programs. Spending on mandatory programs is determined by the number of eligible recipients. 

Discretionary spending must be appropriated each year by the House and Senate. The largest portion of discretionary spending is funding for the Department of Defense and other defense-related programs. Non-defense discretionary spending, just under half of total discretionary spending in 2018, accounts for all federal spending on diplomacy, education, transportation, research, law enforcement, and paying for the operations of the federal government. 

```{r Share of Spending, echo=FALSE}
type_over_time <- budget_longform %>% select(fiscal_year, `BEA Category`, expenditure) %>% filter(fiscal_year != 'TQ') %>% group_by(fiscal_year, `BEA Category`) %>% tally(expenditure) 

#colnames(type_over_time) <- c('fiscal_year', 'discretionary', 'mandatory', 'net_interest')
#type_over_time <- type_over_time %>% mutate(total_spend = discretionary + mandatory + net_interest)
#type_over_time <- type_over_time %>% mutate(prop_disc = discretionary/total_spend)
#type_over_time <- type_over_time %>% mutate(prop_mand = mandatory/total_spend)
#type_over_time <- type_over_time %>% mutate(prop_int = net_interest/total_spend)

ShareOfSpending <- type_over_time %>% group_by(fiscal_year) %>% ggplot(aes(x=fiscal_year, y=n, color=`BEA Category`, fill=`BEA Category`)) +
    geom_area(position = "fill", aes(group=`BEA Category`), color = "black") + 
  scale_y_continuous(expand = c(0, 0), labels = scales::percent_format(), breaks = seq(0,1,by=.2)) + 
    geom_line(color = "black", position = "fill", aes(group=`BEA Category`)) + 
    scale_x_discrete(expand = c(0,0), breaks=(seq(1962,2018,2)), 
                     labels = (seq(1962,2018,2))) +
    scale_fill_brewer(palette="Set2") + 
  
  labs(title = "Mandatory Spending Has Accounted for a \n Larger and Larger Share of the Federal Budget",
         subtitle = "",
         caption = "Source: United States. Office of Management and Budget. Public Budget Database.",
         x = "Year",
         y = "Proportion of Federal Budget") +  
  
    theme(plot.title = element_text(color="black", size=18, face="bold", hjust = 0.5),
          axis.title.x = element_text(color="black", size=18), axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
          axis.title.y = element_text(color="black", size=18), axis.text.y = element_text(size = 12),
          legend.background = element_rect(fill="lightblue",
                                         size=0.5, linetype="solid", 
                                         colour ="black"),
          panel.background = element_rect(fill = NA, size = .2),
          panel.ontop = TRUE)

ggsave(filename = 'visuals/ShareOfSpending.pdf', width = 11, height = 8.5)
ShareOfSpending
```


# Revenues
```{r Revenues Over Time, echo=FALSE}
rev_colnames <- c("Category Code", "Category Name", "Subcategory Code", "Subcategory Name", "Agency Code", "Agency Name", "Bureau Code", "Bureau Name", "Account Code", "Account Name", "Treasury Agency Code", "On/Off Budget")

colnames(rev)[1:12] <- rev_colnames

rev_long <- gather(rev, fiscal_year, expenditure, '1962':'2023', factor_key=TRUE) %>% select("Category Name", "Subcategory Name", "Agency Name", "Bureau Name", "Account Name", "Treasury Agency Code", "On/Off Budget", "fiscal_year", "expenditure")

theme2 <- theme(plot.title = element_text(color="black", size=18, face="bold", hjust = 0.5),
           axis.title.x = element_text(color="black", size=18), axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
           axis.title.y = element_text(color="black", size=18), axis.text.y = element_text(size = 12))

temp <- rev_long %>% group_by(`Category Name`, fiscal_year) %>% select(fiscal_year, `Category Name`, expenditure) %>%
  summarize(sum=sum(expenditure))
temp2 <- filter(temp, `Category Name` != 'Legislative Proposals' & fiscal_year != 'TQ' & !fiscal_year %in% c(2019, 2020, 2021, 2022, 2023))

labsrev <-   labs(title = "Something Interesting About Federal Government Revenues",
         subtitle = "",
         caption = "Source: United States. Office of Management and Budget. Public Budget Database.",
         x = "Fiscal Year",
         y = "Proportion of Federal Revenues")

themerev <- theme(plot.title = element_text(color="black", size=18, face="bold", hjust = 0.5),
                  axis.title.x = element_text(color="black", size=18), axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
                  axis.title.y = element_text(color="black", size=18), axis.text.y = element_text(size = 12),
                  legend.background = element_rect(fill="lightblue",
                                                   size=0.5, linetype="solid", 
                                                   colour ="black"),
                  panel.background = element_rect(fill = NA, size = .2),
                  panel.ontop = TRUE)

rev_share <- temp2 %>% 
  ggplot(aes(x=fiscal_year, y=`sum`, color=`Category Name`, fill=`Category Name`)) +
  geom_area(position = "fill", aes(group=`Category Name`), color = "black") + 
  scale_y_continuous(expand = c(0, 0), labels = scales::percent_format(), breaks = seq(0,1,by=.2)) + 
  geom_line(color = "black", position = "fill", aes(group=`Category Name`)) + 
  scale_x_discrete(expand = c(0,0), breaks=(seq(1962,2023,4)), 
                   labels = (seq(1962,2023,4))) +
  scale_fill_brewer(palette="Set2") + 
  labsrev +  
  themerev 

ggsave(filename = 'visuals/ShareOfRevenue.pdf', width = 11, height = 8.5)
ggsave(filename = 'visuals/ShareOfRevenue.png', width = 11, height = 8.5)


rev_share
```

