---
title: "Geospatial[Redo]"
author: "Ian Lyons"
date: "4/14/2019"
output: 
  html_document: 
    code_folding: hide
    fig_caption: yes
    fig_height: 10
    fig_width: 12
    toc_depth: 1
    toc_float:
      collapsed: false
    theme: cosmo
    toc: yes
urlcolor: blue
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(rmarkdown)
library(here)
library(gghighlight)
library(ggpubr)
library(ggrepel)
library(formatR)
library(tinytex)
library(scales)
library(ggthemes)
library(extrafont)
library(RColorBrewer)
library(treemapify)
library(ggplot2)

library(fiftystater)
library(sf)
library(tools)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r Read in Data & Reshape, message=FALSE, warning=FALSE}
### Define Theme
custom_pal <- c("#3FC1C9", "#B9F3FF", "#8A4DCC", "#4DCC74", "#EEB9FF", "#CC534D", "#CC7776", "#FFF07A")

iltheme <- theme(plot.title = element_text(color="black", 
                                           size=20, 
                                           family="Verdana", 
                                           hjust=0),
                text = element_text(family="Courier New", size=14),
                plot.subtitle = element_text(color="#4c4e4d", 
                                             size=16, 
                                             hjust=0, 
                                             family = "Verdana"),
                axis.title.x = element_text(color="black", size=20, family="Verdana"),
                axis.title.y = element_text(color="black", size=20, family="Verdana"), 
                plot.caption = element_text(color="black", size=14, face="italic"),
                axis.text.x = element_text(size=14),
                axis.text.y = element_text(size=14),
                plot.background=element_rect(fill="white"),
                panel.background =element_rect(fill="white"),
                panel.grid.minor.y = element_blank(),
                axis.ticks = element_blank(),
                axis.line = element_line(size=1, color = "Black"),
                legend.text = element_text(size=16))
```


# Federal Aid to State & Local Governments

In Fiscal Year 2018, the federal government was projected to grant about $728 billion to state and local governments. According to the Congressional Research Service, federal grants make up one-third of state spending, and more than half of state spending on health care and other public assistance programs. Since the expansion of Medicaid included in the Affordable Care Act (ACA), the amount of federal grant money flowing to health care has only increased.

```{r Load in the map of the US, warning=FALSE, fig.height=8, fig.width=9}
raw_aid <- read_csv("../Raw Data/state aid - Sheet1.csv")
library("ggplot2")
theme_set(theme_bw())
library("sf")

library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale='medium',returnclass = 'sf')
usa <- subset(world, admin == "United States of America")


alaska <- ggplot(data = usa) +
     geom_sf(fill = '#cfd0e7') +
     coord_sf(crs = st_crs(3467), xlim = c(-2400000, 1600000), ylim = c(200000, 
         2500000), expand = FALSE, datum = NA)

hawaii  <- ggplot(data = usa) +
     geom_sf(fill='#1d77a8') +
     coord_sf(crs = st_crs(4135), xlim = c(-161, -154), ylim = c(18, 
         23), expand = FALSE, datum = NA)

states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
states <- cbind(states, st_coordinates(st_centroid(states)))
states$ID <- toTitleCase(states$ID)

# Join aid data to states df
joined <- right_join(states, raw_aid, by=c("ID" = "NAME"))
breaks <- quantile(x = joined$`2018 grant per capita`,seq(0,1,.250))
bins <- cut(joined$`2018 grant per capita`, unique(breaks), include.lowest=TRUE)
joined$bin <- bins
bin_labels <- c("$1,450", "$1,660", "$2,020", "$2,460")
joined <- filter(joined, STUSPS != 'DC')

main <- (ggplot(data = usa) +
     geom_sf() +
    geom_sf(data = joined, aes(fill = `2018 grant per capita`)) + 
     coord_sf(crs = st_crs(2163), xlim = c(-2500000, 2500000), ylim = c(-2300000, 
         730000))) +
  scale_fill_distiller(palette = "PuBu", direction = 1, labels = bin_labels, name="Per Capita\nFederal Grant") +
   annotation_custom(
      grob = ggplotGrob(alaska),
      xmin = -2750000,
      xmax = -2750000 + (1600000 - (-2400000))/2.5,
      ymin = -2450000,
      ymax = -2450000 + (2500000 - 200000)/2.5
  ) +
  annotation_custom(
      grob = ggplotGrob(hawaii),
      xmin = -1250000,
      xmax = -1250000 + (-154 - (-161))*120000,
      ymin = -2450000,
      ymax = -2450000 + (23 - 18)*120000
  ) +   
  theme(legend.position = c(.08, .8), 
        legend.background = element_blank(),
        plot.title = element_text(size = 22)) +
  labs(title = "Federal Grant Dollars Flow to Health Care Programs")

main
```

```{r fig.height=8, fig.width=9}
expanded <- c("AK", "HI", "NM", "AZ", "CA", "CO", "NV", "OR", "WA", "MT", "ND", "MN", "AR", "LA", "IA", "MI", "IL", "IN", "OH", "KY", "WV", "DC", "MD", "DE", "CT", "NJ", "PA", "CT", "RI", "MA", "NH", "VT", "NY")
joined <- joined %>% mutate('Expanded' = case_when((STUSPS %in% expanded) ~ 'Expanded Medicaid', (!STUSPS %in% expanded) ~ 'Did Not Expand Medicaid'))

column_pc <- ggplot(data = joined, 
                    aes(x = reorder(STUSPS, `2018 grant per capita`), 
                        y = `2018 grant per capita`)) + 
  geom_col(data = joined, 
           aes(fill = Expanded, 
               label = STUSPS),
           color="black", 
           size=0.25) +
  iltheme +  
  scale_y_continuous(name="Grant Amount ($US) Per Person", 
                     labels = scales::dollar_format(),
                     expand = c(0,0)) +
  scale_x_discrete(name= element_blank(), 
                   expand = expand_scale()) +
  scale_fill_brewer(palette = "Blues", name="Expanded Medicaid?") +
  theme(plot.title = element_text(size=22),
        panel.grid = element_blank(), 
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(colour = "#AFAFAF"),
        axis.text.x = element_text(size=12, 
                                   hjust = 1),
        axis.text.y = element_text(size=12),
        legend.background = element_rect(size = .25,  
                                         linetype="solid", 
                                         colour ="black"),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18, 
                                    face = "bold"),
        panel.background = element_rect(fill = NA, 
                                        size = .2),
        legend.position = c(.75, .2),
        axis.line = element_line(size = .25)) + 
  labs(title = "Federal Grants Support Medicaid Expansion") +
  coord_flip()
column_pc 

```



